---
title: "rcoco"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
vignette: >
  %\VignetteIndexEntry{Quick introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, cache = FALSE}
#library(mlrMBO)
set.seed(123)
knitr::opts_chunk$set(cache = TRUE, collapse = FALSE)
knitr::knit_hooks$set(document = function(x){
  gsub("```\n*```r*\n*", "", x)
})
```

# Quickstart

The [COCO framework](https://github.com/numbbo/coco) (COmparing Continuous Optimisers) allows for benchmarking continuous black-box optimizers on the BBOB (Black-Box-Optimization-Benchmarking) functions. Its post-processing module allows standardized evaluation and visualization of optimizer results. The R package *rcoco* is an connector to the C core of the COCO framework. It allows to benchmark optimizers written in R. In the following we describe the workflow by example: we run the Nelder-Mead simplex algorithm and BFGS (Broyden-Fletcher-Goldfarb-Shanno) on the noiseless BBOB test set.

We need to perform the following steps:

1. Generate the benchmark suite, i.e., a set of test functions.
1. Implement a simple wrapper for each optimizer.
1. Create an observer for each optimizer and run each optimizer on the benchmark suite.
1. Postprocess the output data, i.e, generate standardized evaluation(s) and visualizations.

## Initializing a benchmark suite

The first step is to initialize a benchmark suite. This is basically a collection of benchmarking functions with a fixed number of objectives. Here we want to benchmark both algorithms on a subset of the single-objective noiseless BBOB functions. In the COCO framework functions are described by an integer identifier (ID) and further varied by decision space dimension and another parameter responsible, e.g., for rotation and shifting of the function (see the [official documentation](http://numbbo.github.io/coco-doc/#functions-instances-and-problems) for further details).
A suite with the BBOB functions with IDs 10, 11 and 12 with n = 2, 3 dimensions and three variations (1, 2, 3) are created as follows in rcoco:

```{r}
library(rcoco)
suite = cocoOpenSuite("bbob", instances = 10:12, dims = c(2, 3), inst.inds = 1:3)
print(suite)
```

## Preparing the optimizers

Next step on the ToDo list: implement a simple wrapper function for each optimizer under consideration. A wrapper is a function which expects the objective function `fn`, a COCO problem `problem` and additional arguments which may be passed down the optimizer. Here, the wrapper functions simply delegate the objective functions and parameters extracted from the COCO problem to the `optim` function from the `stats` package:

```{r}
optNM = function(fn, problem, ...) {
  optim(par = problem$initial.solution, fn = fn, method = "Nelder-Mead", ...)
}
optBFGS = function(fn, problem, ...) {
  optim(par = problem$initial.solution, fn = fn, method = "BFGS", ...)
}
```

## Run optimizer on suite

Up next we need to setup an observer for each optimzer and apply each optimzer to the suite. A COCO observer logs function evaluations and stores these alongside further meta information in a specific folder structure. Here we pass the algorithm name. Hence, the data will be stored in the folder *expdata/cocoexp/algorithmname*.

```{r, echo=-(1:3)}
# we need to do this again here since pointers cannot be cached
cocoCloseSuite(suite)
suite = cocoOpenSuite("bbob", instances = 10:12, dims = c(2, 3), inst.inds = 1:3)
observer = cocoInitObserver(suite, algorithm.name = "NelderMead")
res = cocoSuiteRunOptimizer(suite, optNM, observer)

observer = cocoInitObserver(suite, algorithm.name = "BFGS")
res = cocoSuiteRunOptimizer(suite, optBFGS, observer)

cocoCloseSuite(suite)
```

## Postprocessing

Now, the output data is stored in the folder *exdata/cocoexp/{NelderMead,BFGS}*. The final step is the statistical analysis. COCO offers a sophisticated post-processing module which produces a plethora of visualizations and tables. This step needs to be performed from the outside of R. Open a bash or terminal, navigate to the working directory of your R session and type

```
#python -m cocopp exdata/cocoexp/NelderMead exdata/cocoexp/BFGS
```

The post-processing may take some minutes depending on the number of optimizers, their parameterization and the size of the benchmark suite. Go grab some coffee or cookies to fill the time. After a while the output from post-processing will be available in the *ppdata* folder by default. Open *ppdata/index.html* for convenient exploration of the post-processing results in your favourite webbrowser (see the [official COCO documentation](https://github.com/numbbo/coco) for additional information on post-processing capabilities).
